### 2022-01-07 content/80.formal-definition.tex
URIs -> IRIs
typo fixes as in proof

\begin{eqnarray*}
    \mathcal{L}_{rocrate}   & = & \big\{ Property(p), Class(c),
                            Value(x), \mathbb{R}, \mathbb{S} \big\} \\
    \mathbb{D}              & = & \mathbb{IRI} \\
    \mathbb{IRI}            & \equiv & { \text{IRIs as defined in RFC3987} } \\
    \mathbb{R}              & \equiv & { \text{real or integer numbers} } \\
    \mathbb{S}              & \equiv & { \text{literal strings} }
\end{eqnarray*}

The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers \cite{doi:10.17487/rfc3987} (notation $\texttt{<http://example.com/>}$)\footnote{
    For simplicity, blank nodes are not included in this formalisation, as RO-Crate
    recommends the use of IRI identifiers: \url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#describing-entities-in-json-ld}
}, with additional descriptions using numbers $\mathbb{R}$ (notation $13.37$) and literal strings $\mathbb{S}$ (notation $\text{“Hello”}$).

@@

An $ROCrate(R)$ is defined as a self-described \emph{Root Data Entity}, which describes and contains parts (\emph{data entities}), which are further described in \emph{contextual entities}.  These terms align with their use in the RO-Crate 1.1 terminology\footnote{
    \url{https://www.researchobject.org/ro-crate/1.1/terminology}}.

The $Root(r)$ is a type of $Dataset(r)$, and must have the metadata to literal attributes to provide a $name$, $description$ and $datePublished$, as well as a contextual entity identifying its license. These predicates correspond to the RO-Crate 1.1 requirements for the root data entity\footnote{
    \url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}
}.

@@

The below is an example RO-Crate represented using the above formalisation, assuming a base URI of \texttt{<http://example.com/ro/123/>}:

\allowdisplaybreaks
\begin{eqnarray*}
&& ROCrate(\texttt{<http://example.com/ro/123/>}) \\
&& name(\texttt{<http://example.com/ro/123/>}, \\
&& \ \ \ \ \ \text{“Data files associated with the manuscript:Effects of …”}) \\
&& description(\texttt{<http://example.com/ro/123/}, \\
&& \ \ \ \ \ \text{“Palliative care planning for nursing home residents …”}) \\
&& license(\texttt{<http://example.com/ro/123/>}, \\
&& \ \ \ \ \ \texttt{<https://spdx.org/licenses/CC-BY-4.0>}) \\
&& datePublished(\texttt{<http://example.com/ro/123/>}, \text{“2017-02-23”}) \\
&& hasPart(\texttt{<http://example.com/ro/123/>},
        \texttt{<http://example.com/ro/123/file.txt>}) \\
&& hasPart(\texttt{<http://example.com/ro/123/>},
        \texttt{<http://example.com/ro/123/interviews/>}) \\
\\
&& ContextualEntity(\texttt{<https://spdx.org/licenses/CC-BY-4.0>}) \\
&& name(\texttt{<https://spdx.org/licenses/CC-BY-4.0>},  \\
&& \ \ \ \ \  \text{Creative Commons Attribution 4.0”}) \\
\\
&& ContextualEntity(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>}) \\
&& name(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>},  \\
&& \ \ \ \ \  \text{Creative Commons Attribution Non Commercial 4.0”}) \\
\\
&& File(\texttt{<http://example.com/ro/123/survey.csv>}) \\
&& name(\texttt{<http://example.com/ro/123/survey.csv>},
        \text{“Survey of care providers”}) \\
\\
&& Dataset(\texttt{<http://example.com/ro/123/interviews/>}) \\
&& name(\texttt{<http://example.com/ro/123/interviews/>},  \\
&& \ \ \ \ \  \text{“Audio recordings of care provider interviews”}) \\
&& license(\texttt{<http://example.com/ro/123/interviews/>}, \\
&& \ \ \ \ \ \texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>})
\end{eqnarray*}

@@

A formalized RO-Crate in $\mathcal{L}_{rocrate}$ can be mapped to different serializations.
Assume a simplified\footnote{
 This simplification does not cover the extensive list of literal datatypes built-in to RDF 1.1, only strings and decimal real numbers. Likewise, language of literals are not included.
} language $\mathcal{L}_{RDF}$
based on the RDF abstract syntax \cite{rdfworkinggroup_2014}:

@@

An important RO-Crate principle is that of being \textbf{self-described} Therefore the serialisation of the RO-Crate into a file should also describe itself in a Metadata File Descriptor\footnote{
    \url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#ro-crate-metadata-file-descriptor}
}, indicating it is $about$ (describing) the RO-Crate root data entity, and that it $conformsTo$ a particular version of the RO-Crate specification:

>>>>

\begin{eqnarray*}
    \mathcal{L}_{rocrate}   & = & \big\{ Property(p), Class(c),
                            Value(x), \mathbb{R}, \mathbb{S} \big\} \\
    \mathbb{D}              & = & \mathbb{IRI} \\
    \mathbb{IRI}            & \equiv & { \text{IRIs as defined in RFC3987 \cite{doi:10.17487/rfc3987}}} } \\
    \mathbb{R}              & \equiv & { \text{real or integer numbers} } \\
    \mathbb{S}              & \equiv & { \text{literal strings} }
\end{eqnarray*}

The domain of discourse $\mathbb{D}$ is the set of $\mathbb{IRI}$ identifiers (notation $\texttt{<http://example.com/>}$)\footnote{
    For simplicity, blank nodes are not included in this formalisation, as RO-Crate
    recommends the use of IRI identifiers: \url{https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html\#describing-entities-in-json-ld}
}, with additional descriptions using numbers $\mathbb{R}$ (notation $13.37$) and literal strings $\mathbb{S}$ (notation $\text{“Hello”}$).

@@

An $ROCrate(R)$ is defined as a self-described \emph{Root Data Entity}, which contains parts (\emph{data entities}), which are further described in \emph{contextual entities}.  These terms align with their use in the RO-Crate 1.1 terminology\footnote{
    \url{https://www.researchobject.org/ro-crate/1.1/terminology}}.

The $Root(r)$ is a type of $Dataset(r)$, and must as metadata have at least the attributes $name$, $description$ and $datePublished$, as well as a contextual entity that identify its $license$. These predicates correspond to the RO-Crate 1.1 minimal requirements for the root data entity\footnote{
    \url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#direct-properties-of-the-root-data-entity}
}.

@@

The below is an example RO-Crate represented using the above formalisation, assuming a base IRI of \texttt{<http://example.com/ro/123/>}:

\allowdisplaybreaks
\begin{eqnarray*}
&& ROCrate(\texttt{<http://example.com/ro/123/>}) \\
&& name(\texttt{<http://example.com/ro/123/>}, \\
&& \ \ \ \ \ \text{“Data files associated with the manuscript:Effects of …”}) \\
&& description(\texttt{<http://example.com/ro/123/}, \\
&& \ \ \ \ \ \text{“Palliative care planning for nursing home residents …”}) \\
&& license(\texttt{<http://example.com/ro/123/>}, \\
&& \ \ \ \ \ \texttt{<https://spdx.org/licenses/CC-BY-4.0>}) \\
&& datePublished(\texttt{<http://example.com/ro/123/>}, \text{“2017-02-23”}) \\
&& hasPart(\texttt{<http://example.com/ro/123/>},
        \texttt{<http://example.com/ro/123/file.txt>}) \\
&& hasPart(\texttt{<http://example.com/ro/123/>},
        \texttt{<http://example.com/ro/123/interviews/>}) \\
\\
&& ContextualEntity(\texttt{<https://spdx.org/licenses/CC-BY-4.0>}) \\
&& name(\texttt{<https://spdx.org/licenses/CC-BY-4.0>},  \\
&& \ \ \ \ \  \text{“Creative Commons Attribution 4.0”}) \\
\\
&& ContextualEntity(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>}) \\
&& name(\texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>},  \\
&& \ \ \ \ \  \text{“Creative Commons Attribution Non Commercial 4.0”}) \\
\\
&& File(\texttt{<http://example.com/ro/123/survey.csv>}) \\
&& name(\texttt{<http://example.com/ro/123/survey.csv>},
        \text{“Survey of care providers”}) \\
\\
&& Dataset(\texttt{<http://example.com/ro/123/interviews/>}) \\
&& name(\texttt{<http://example.com/ro/123/interviews/>},  \\
&& \ \ \ \ \  \text{“Audio recordings of care provider interviews”}) \\
&& license(\texttt{<http://example.com/ro/123/interviews/>}, \\
&& \ \ \ \ \ \texttt{<https://spdx.org/licenses/CC-BY-NC-4.0>})
\end{eqnarray*}

@@

A formalized RO-Crate in $\mathcal{L}_{rocrate}$ can be mapped to different serializations.
Assume a simplified\footnote{
 This simplification and mapping does not cover the extensive list of literal datatypes built into RDF 1.1, only strings and decimal real numbers. Likewise, $LanguageTag$ is deliberately not utillised below.
} language $\mathcal{L}_{RDF}$
based on the RDF abstract syntax \cite{rdfworkinggroup_2014}:

@@

An important RO-Crate principle is that of being \textbf{self-described}. Therefore the serialisation of the RO-Crate into a file should also describe itself in a Metadata File Descriptor\footnote{
    \url{https://www.researchobject.org/ro-crate/1.1/root-data-entity.html\#ro-crate-metadata-file-descriptor}
}, indicating it is $about$ (describing) the RO-Crate root data entity, and that it $conformsTo$ a particular version of the RO-Crate specification:
